Direct comparison between simulations and observations: radio emission separated into star formation and AGN activity

### 2024-08-21 src/tex/ms.tex
[showyourwork] overleaf sync

\newcommand{\sarc}{$^{\prime\prime}\!\!$}
\newcommand{\myreferences}{references.bib}
\newcommand{\frii}{FR$\,$II}
\newcommand{\fri}{FR$\,$I}
\newcommand{\wphz}{$\,$W$\,$Hz$^{-1}$}
\newcommand{\tbpeak}{$T_{b,\mathrm{peak}}$}
\newcommand{\tbtotal}{$T_{b,\mathrm{total}}$}
\newcommand{\muJybm}{$\mu$Jy$\,$beam$^{-1}$}

@@

% Title
\title[TBD]{TBD}

@@

\subsection{Simulations}
\hl{Nicole}

\section{Methods}

@@

In Fig.~\ref{fig:agn_evolution} we show

@@

%\begin{figure}
%    \centering
%    \includegraphics[width=0.5\textwidth]{figures/simba_comparison.png}
%    \caption{Caption}
%    \label{fig:simba}
%    \script{simba_comparison.py}
%\end{figure}

\begin{figure*}
    \centering
    \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_SF.png}
    \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_SF.png}
    \caption{Caption}
    \label{fig:sf_evolution}
    \script{RLF_evolution_v2.py}
\end{figure*}

\section{Discussion and conclusions}

>>>>

\newcommand{\sarc}{$^{\prime\prime}\!\!$}
\newcommand{\myreferences}{references.bib}
\newcommand{\frii}{FR$\,$II}
\newcommand{\fri}{FR$\,$I}
\newcommand{\wphz}{$\,$W$\,$Hz$^{-1}$}
\newcommand{\tbpeak}{$T_{b,\mathrm{peak}}$}
\newcommand{\tbtotal}{$T_{b,\mathrm{total}}$}
\newcommand{\muJybm}{$\mu$Jy$\,$beam$^{-1}$}
\newcommand{\llof}{$L_{\textrm{144MHz}}$}

@@

% Title
\title[TBD]{Direct comparison between simulations and observations: radio emission separated into star formation and AGN activity}

@@

\subsection{Simulations}
\label{subsec:simba}
\hl{Nicole}\\
We use the SIMBA cosmological hydrodynamic simulation which accounts for an H$_{2}$ based star formation rate following the sub-grid model of \hl{Krumholtz $\&$ Gnedin (2011)}. Additionally, SIMBA models the growth of supermassive black holes via a two-mode prescription based on the temperature of gas within a distance, $R_{0}$, calculated over the 256 nearest particles limited to 2$h^{-1}$kpc. 
It is commonplace for cosmological simulations to employ the growth of SMBHs via the Bondi rate \hl{citations EAGLE, HORIZON AGN, Illustris, TNG}. In SIMBA, the Bondi rate is employed when gas within $R_{0}$ has temperatures T$>10^{5}$K, referred to as the hot mode. What makes SIMBA unique is the additional contribution to the accretion rate via a torque limited accretion flow when gas has temperatures T$<10^{5}$K, or in the cold mode. 
SIMBA also employs a physically motivated model for AGN feedback in the form of AGN winds \hl{cite Perna}, as well as AGN jets with velocities that increase inversely proportional to the Eddington rate in accordance with \hl{Heckman and Best 2014}.
SIMBA therefore provides a platform to study jet-mode AGN in a way that has not been explored before. 
Specifically, we select galaxies in SIMBA that have full-velocity jets, that is, when the Eddington rate drops below 0.02 (but non-zero) and where the black hole mass is larger than M$_{\rm BH}>10^{8}$M$_{\odot}$. For this population we are able to calculate the amount of radio emission due to star formation processes following \hl{Condon 1992} as well as that from accretion processes following \hl{Kording, Jester and Fender 2008}. Details of this method can be found in \hl{Thomas2021}. We can therefore calculate the contribution to the radio luminosity function from star formation and AGN accretion processes. We can additionally separate the galaxy population into star forming and quiescent galaxies, radio-excess, or radiatively efficient and inefficient, based on the sources black hole and host properties.
\section{Methods}

@@

In Fig.~\ref{fig:evolution} we show a comparison between the observed RLFs and those generated from SIMBA (see \S~\ref{subsec:simba}), and the redshift evolution for both. The left set of panels shows the results for SF and the right set of panels shows the results for AGN. \hl{We cut off the RLFs below a flux limit where they are not constrained}. We use the same redshift bins across all samples for consistency. We reproduce previous observational results that while the SF population has a strong redshift evolution, the AGN population does not. For the simulations, the behaviour is qualitatively similar, although the AGN population does have mild redshift evolution whereas the observations do not.

The bottom panels in Fig~\ref{fig:evolution} show the ratio of activity to galaxy RLFs, the same as in the bottom right hand panel of Fig~\ref{fig:rlfs}. Comparing these ratios between observations and simulations, we see qualitatively the same behaviour, although the simulations reach much lower and higher values for both the SF and AGN populations, respectively. In the lowest redshift bin, the ratio for the simulation stabilises to around unity for \llof $<10^{23}\,$\wphz , and decreases rapidly above this, reaching much smaller values than the observations. Higher redshift bins probe successively higher luminosity ranges, also reaching small values for the ratio between activity and galaxies. This is consistent with the fact that at higher redshifts, most radio AGN in SIMBA are hosted in SFGs, and thus the \textit{galaxy} RLF for SIMBA SFGs is inflated with AGN emission.

For the AGN population (Fig~\ref{fig:evolution}, right set of panels), the ratio between observed activity and galaxy RLFs  converges to unity above $\sim$\llof $10^{24}\,$\wphz\ for the lowest redshift bin. This convergence happens at higher luminosity values for higher redshifts. Below these luminosities, across all redshift bins, the ratio increases. The maximum value for each redshift bin approaches $\sim$2, although the fact that in the lowest redshift bin the value stabilises at $\sim1.3$ and only climbs higher at the very lowest luminosities (\llof $<10^{22}\,$\wphz ) may be due to incompleteness. If we disregard the lowest luminosity bins, it seems that we see a 30 - 50 percent increase in AGN activity over what is measured when using galaxy classifications alone, for all redshifts. The ratio for the SIMBA RLFs shows strong redshift evolution, and much higher ratio values. The RLFs themselves turn over, whereas the observed RLFs do not. This is likely a consequence of the fact that the AGN emission in SIMBA is due to high-velocity jets, whereas there is increasing observational evidence for low-luminosity jets or AGN emission \citep{macfarlane_radio_2021,yue_novel_2024}. The high values reached by the ratio of RLFs again is expected as many high-$z$ radio AGN in SIMBA are hosted in SFGs. Shifting this AGN emission to AGN \textit{activity} when it is counted in the SFG \textit{galaxy} population generates high values for the activity to galaxy ratio.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_SF.png}
    \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_AGN.png}
    \caption{Caption}
    \label{fig:evolution}
    \script{RLF_evolution_v2.py}
\end{figure*}

\section{Discussion}

\section{Conclusions}
